---
title: "Introduction à Shiny"
author: "Antoine Neuraz - antoine.neuraz@gmail.com"
date: "18/10/2016"
output:
  ioslides_presentation:
    highlight: pygments
    keep_md: yes
    widescreen: yes
runtime: shiny
---

```{r setup, include=FALSE}
library(shiny)
library(dplyr)
library(ggplot2)
library(plotly)
library(DT)
library(shinydashboard)

knitr::opts_chunk$set(echo = FALSE)

catalog_path <- 'data/phewas-catalog.csv'
codes_path <- 'data/phewas_codes_2016-10-11.csv'

# load data from csv
catalog <- read.csv(catalog_path, stringsAsFactors = F)
codes <- read.csv(codes_path, stringsAsFactors = F)

# extract unique phewas codes from codes list
codes <- codes %>% 
  select(jd_code, category_string) %>%
  unique()

# merge codes with their respective category (for plotting)
catalog <- catalog %>% merge(codes, by.x = 'phewas.code', by.y = 'jd_code')

# extract the list of snps from the catalog
snps <- catalog %>% 
  select(snp, gene_name) %>%
  unique() 

# select top10 genes with significant associations 
top10 <- catalog %>% 
  filter(p.value < 0.05/1815) %>% 
  group_by(snp, gene_name) %>%
  summarize(count = n()) %>% 
  arrange(desc(count)) %>%
  head(10) %>% 
  select(snp, gene_name)


# function to draw a manhattan plot with ggplot2 
# parameters: catalog = dataframe with the results from the PheWAS catalog
#             choice = a named vector with the snp and the gene_name
draw_manhattan_simple <- function(catalog, choice) {

  snp_sel <- choice$snp
  gene_sel <- choice$gene_name

  
}

draw_manhattan_plotly <- function(catalog, choice) {
  
}

```

## Example simple d'application Shiny

```{r eruptions, echo = F}

shinyApp(
  
  ui <- fluidPage(
    inputPanel(
      selectInput("n_breaks", label = "Nombre de barres:",
              choices = c(10, 20, 35, 50), selected = 20),
      sliderInput("bw_adjust", label = "Largeur d'ajustement pour la densité:",
              min = 0.2, max = 2, value = 1, step = 0.2)
      ),
    plotOutput("geyserPlot")
    ),
  
  server <- function(input, output) {
   output$geyserPlot  <- renderPlot({
      hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
           xlab = "Durée (minutes)", main = "Durée d'éruption du Geyser")
      dens <- density(faithful$eruptions, adjust = input$bw_adjust)
      lines(dens, col = "blue")
      }) 
  })

```

## Récupérer les données

- [PheWAS Catalog](https://phewas.mc.vanderbilt.edu/datatable)
- [PheWAS codes](https://phewas.mc.vanderbilt.edu/phecodes/DB/get_all_jdcodes.html) 

### cloner le repo Github du cours
- [https://github.com/aneuraz/shiny-intro](https://github.com/aneuraz/shiny-intro)

## Shiny, qu'est-ce que c'est ? 
- package R développé par RStudio. 
- web-applications dynamiques
- HTML, css, javascript FREE


## Anatomie d'une app shiny

```{r anatomy, echo = T}
shinyApp(
  
  ui <- basicPage(),
  
  server <- function(input, output) {}
  
)
```

## Anatomie d'une app shiny: ui 

- powered by Twitter Bootstrap

```{r ui, echo = T}

shinyApp(
  
  ### <b>
  ui <- basicPage( h2("texte affiché directement par l'ui") ),
  ### </b>
  
  server <- function(input, output) {}
  
)

```

## Anatomie d'une app shiny: server

```{r server, echo = T}

shinyApp(
  
  ui <- basicPage( h2("texte affiché directement par l'ui"),
                   h2(textOutput("texte_serveur"))),
  
  ### <b>
  server <- function(input, output) {
    output$texte_serveur <- renderText("texte affiché via le server")
  }
  ### </b>
)

```

## Layouts: sidebar

```{r ui-layout1, echo = T, eval = F}

shinyApp(
  
  ui <- fluidPage( 
          titlePanel("Ceci est le titre"),
          
          sidebarLayout(
            sidebarPanel(
              "Sidebar"
              ),
          
            mainPanel(
              h2(textOutput("texte_serveur")),
              tags$blockquote("Voici une citation...")
            )
          )
          ),
  
  ### <b>
  server <- function(input, output) {
    output$texte_serveur <- renderText("texte affiché via le server")
  }
  ### </b>
)

```


## Layouts: sidebar-app

```{r ui-layout1-app, echo = F, eval = T}

shinyApp(
  
  ui <- fluidPage( 
          titlePanel("Ceci est le titre"),
          
          sidebarLayout(
            sidebarPanel(
              "Sidebar"
              ),
          
            mainPanel(
              h2(textOutput("texte_serveur")),
               tags$blockquote("Voici une citation...")
            )
          )
          ),
  
  ### <b>
  server <- function(input, output) {
    output$texte_serveur <- renderText("texte affiché via le server")
  }
  ### </b>
)

```

## Layouts: tabset

```{r ui-layout3, echo = T, eval = F}

shinyApp(
  
  ui <- fluidPage( 
    titlePanel("Ceci est le titre"),
    tabsetPanel(
      tabPanel("Plot", textOutput("texte_plot")),
      tabPanel("Table", textOutput("texte_table"))
    )
  ),
  
  ### <b>
  server <- function(input, output) {
    output$texte_plot <- renderText("ceci est un plot")
    output$texte_table <- renderText("ceci est un tableau")
  }
  ### </b>
)
```

## Layouts: tabset-app

```{r ui-layout4, echo = F, eval = T}

shinyApp(
  
  ui <- fluidPage( 
    titlePanel("Ceci est le titre"),
    tabsetPanel(
      tabPanel("Plot", textOutput("texte_plot")),
      tabPanel("Table", textOutput("texte_table"))
    )
  ),
  
  ### <b>
  server <- function(input, output) {
    output$texte_plot <- renderText("ceci est un plot")
    output$texte_table <- renderText("ceci est un tableau")
  }
  ### </b>
)

```


## Layouts: Grid

```{r ui-layout5, echo = T, eval = F}

shinyApp(
  
  ui <- fluidPage( 
    titlePanel("Ceci est le titre"),
    fluidRow(
      wellPanel(column(4, h2("col1"))),
      column(4, h2("col2")),
      column(4, h2("col3"))
    ),
    fluidRow(
      column(6, h2("col4")),
      column(6, h2("col5"))
    )
  ),
  server <- function(input, output) {
  }
)
```

## Layouts: grid-app

```{r ui-layout6, echo = F, eval = T}

shinyApp(
  
  ui <- fluidPage( 
    titlePanel("Ceci est le titre"),
    fluidRow(
      column(4,  wellPanel(h2("col1"))),
      column(4, h2("col2")),
      column(4, wellPanel(h2("col3")))
    ),
    fluidRow(
      column(6, wellPanel(h2("col4"))),
      column(6, h2("col5"))
    )
  ),
  server <- function(input, output) {
  }
)

```

## Inputs

```{r ui-inputs, echo = T, eval = F}

shinyApp(
  
  ui <- fluidPage( 
    
          sidebarLayout(
            sidebarPanel(
              textInput(inputId = "text_input", label = 'texte')
              ),
          )
          ),
  
  ### <b>
  server <- function(input, output) {
    output$texte_serveur <- renderText("texte affiché via le server")
  }
  ### </b>
)

```

## Inputs-app
```{r ui-inputs-app, echo = F, eval = T}

shinyApp(
  
  ui <- fluidPage( 
    
          sidebarLayout(
            sidebarPanel(
              textInput(inputId = "text_input", label = 'texte'),
              sliderInput(inputId = "slider_input", label = 'slider', min = 0, max = 100, value = 40),
              actionButton(inputId = "button", label="bouton")
              ),
            mainPanel(),
          )
          ),
  
  ### <b>
  server <- function(input, output) {
    output$texte_serveur <- renderText("texte affiché via le server")
  }
  ### </b>
)

```

## Basic Inputs (rstudio shiny tutorial) {.smaller}

```{r input-list, echo = F}
shinyApp(
  ui <- fluidPage(
  fluidRow(
    
    column(3,
      h3("Buttons"),
      actionButton("action", label = "Action"),
      br(),
      br(), 
      submitButton("Submit")),
    
    column(3,
      h3("Single checkbox"),
      checkboxInput("checkbox", label = "Choice A", value = TRUE)),
    
    column(3, 
      checkboxGroupInput("checkGroup", 
        label = h3("Checkbox group"), 
        choices = list("Choice 1" = 1, 
           "Choice 2" = 2, "Choice 3" = 3),
        selected = 1)),
    
    column(3, 
      dateInput("date", 
        label = h3("Date input"), 
        value = "2014-01-01"))   
  ),
  
  fluidRow(
    
    column(3,
      dateRangeInput("dates", label = h3("Date range"))),
    
    column(3,
      fileInput("file", label = h3("File input"))),
    
    column(3, 
      h3("Help text"),
      helpText("Note: help text isn't a true widget,", 
        "but it provides an easy way to add text to",
        "accompany other widgets.")),
    
    column(3, 
      numericInput("num", 
        label = h3("Numeric input"), 
        value = 1))   
  ),
    
  fluidRow(
    
    column(3,
      radioButtons("radio", label = h3("Radio buttons"),
        choices = list("Choice 1" = 1, "Choice 2" = 2,
                       "Choice 3" = 3),selected = 1)),
    
    column(3,
      selectInput("select", label = h3("Select box"), 
        choices = list("Choice 1" = 1, "Choice 2" = 2,
                       "Choice 3" = 3), selected = 1)),
    
    column(3, 
      sliderInput("slider1", label = h3("Sliders"),
        min = 0, max = 100, value = 50),
      sliderInput("slider2", "",
        min = 0, max = 100, value = c(25, 75))
      ),
    
    column(3, 
      textInput("text", label = h3("Text input"), 
        value = "Enter text..."))   
  )

),
server <- function(input, output) {}

)


```

# Interactions

## simple
```{r reactivity_1, echo = T, eval = T}

shinyApp(
  
  ui <- basicPage(
    textInput("inp_text", label = "Texte en entrée"),
    textOutput("out_text")
    ),
  
  server <- function(input, output) {
    output$out_text <- renderText(toupper(input$inp_text))
  }

)

```

## simple - output multiple
```{r reactivity_2, echo = T, eval = T}

shinyApp(
  
  ui <- basicPage(
    textInput("inp_text", label = "Texte en entrée"),
    textOutput("out_text"),
    textOutput("out_num")
    ),
  
  server <- function(input, output) {
    output$out_text <- renderText(toupper(input$inp_text))
    output$out_num <- renderText(as.numeric(input$inp_text))
  }

)

```

## reactive() - no side effect
```{r reactivity_3, echo = T, eval = T}

shinyApp(
ui <- basicPage(
    textInput("inp_text", label = "Texte en entrée"),
    textOutput("out_text")
    ),
  
server <- function(input, output) {
  ### <b>
  reactive_text <- reactive({ 
    input$inp_text %>% toupper() %>% strsplit('') %>% unlist() %>% paste0(collapse = '-')
    })
  ### </b>
  output$out_text <- renderText( reactive_text() )
  }
)

```

## observe() - with side effects
```{r reactivity_4, echo = T, eval = T}

shinyApp(
ui <- basicPage(
    textInput("inp_text", label = "Texte en entrée"),
    textInput("inp_text2", label = "Value:")
    ),
  
server <- function(input, output, session) {
  observe({
    if(input$inp_text == 'bleu') {updateTextInput(session, inputId = "inp_text2", value = "OK!!")}
   })
  }
)

```

## Plotly
```{r manhattan-plotly, echo = T, eval = F}

shinyApp(
  
  ui <- fluidPage(
          fluidRow(
            column(12,
                   plotlyOutput("manhattan")
                   )
          )
          ),
  ### <b>
  server <- function(input, output) {
    p <- catalog %>% ggplot(aes(x=cases)) + geom_histogram()
    output$manhattan <- renderPlotly(ggplotly(p))
  }
  ### </b>
)
```


## Plotly - app
```{r manhattan-plotly2, echo = F, eval = T}

shinyApp(
  
  ui <- fluidPage(
          fluidRow(
            column(12,
                   plotlyOutput("manhattan")
                   )
          )
          ),
  ### <b>
  server <- function(input, output) {
    p <- catalog %>% ggplot(aes(x=cases)) + geom_histogram()
    output$manhattan <- renderPlotly(ggplotly(p))
  }
  ### </b>
)
```


## table with DT
```{r DT, echo = T, eval = F}

shinyApp(
  
  ui <- fluidPage(
          fluidRow(
            column(12,
                   DT::dataTableOutput("cat_table")
                   )
          )
          ),
  ### <b>
  server <- function(input, output) {
    output$cat_table <- DT::renderDataTable(
      catalog %>% filter(snp %in% top10$snp) %>% select(snp, phewas.phenotype, p.value, odds.ratio)
    )
  }
  ### </b>
)
```

## table with DT - app
```{r DT2, echo = F, eval = T}

shinyApp(
  
  ui <- fluidPage(
          fluidRow(
            column(12,
                   DT::dataTableOutput("cat_table")
                   )
          )
          ),
  ### <b>
  server <- function(input, output) {
    output$cat_table <- DT::renderDataTable(
      catalog %>% filter(snp %in% top10$snp) %>% select(snp, phewas.phenotype, p.value, odds.ratio)
    )
  }
  ### </b>
)
```


## References

- [shiny tutorial - rstudio](http://shiny.rstudio.com)
- [R powered web applications with Shiny - zevross.com](http://zevross.com/blog/2016/04/19/r-powered-web-applications-with-shiny-a-tutorial-and-cheat-sheet-with-40-example-apps/)
- [widgets html pour shiny](http://www.htmlwidgets.org)
- [shiny dashboards](https://rstudio.github.io/shinydashboard/get_started.html)

